---
title: "Speaker vs Listener Noise Ratios"
author: "Zach"
date: "2024-01-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(profvis)
#nstall.packages("microbenchmark")
library(microbenchmark)
library(furrr)
library(tictoc)
library(ggpubr)
source('iterated_learning.R')

```

## Speaker vs Listener Noise

In this section we'll manipulate the relationship between speaker (prior_noise) and listener (p_noise).

### p_noise = 0.01, prior_noise = 0.001

We'll keep p_theta equal to 0.5 and prior_mu equal to 0.6.

```{r warning = F, message = F}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 1/100 #listener noise
prior_prob_noise = 1/1000 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.001_listener_.01.csv'

if (!file.exists(file_path)) {

new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.001_listener_.01.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.001_listener_.01.csv')
}

plot_data = new_df 

p_higher_reg = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  theme_bw()

ggarrange(p_lower_reg, p_higher_reg, ncol = 1)
#ggsave('./Write-up/Figures/higher_and_lower_reg.png')
```

### p_noise = 0.001, prior_noise = 0.01

```{r warning = F, message = F}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 1/1000 #listener noise
prior_prob_noise = 1/100 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.001.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.001.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.001.csv')
}

plot_data = new_df 

ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  theme_bw()

#ggsave('./Write-up/Figures/speaker_noise_01_listener_001.png')
```

### p_noise = 0.1, prior_noise = 0.091

Here we keep the relative difference the same

```{r}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.1 #listener noise
prior_prob_noise = 0.091 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.1_listener_.091.csv'

if (!file.exists(file_path)) {

new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.1_listener_.091.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.1_listener_.091.csv')
}

plot_data = new_df 

rel_dif2 = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0, 0.5, 1.0)) +
  theme_bw()


ggarrange(rel_dif2, p_higher_reg, nrow = 2)
#ggsave('./Write-up/Figures/absolute_matters.png')
```

### p_noise = 0.1, prior_noise = 0.01

Here, p_noise is still 10x greater than prior_noise, but the relative difference is different:

```{r}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.1 #listener noise
prior_prob_noise = 1/100 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.1_listener_.01.csv'

if (!file.exists(file_path)) {

new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.1_listener_.01.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.1_listener_.01.csv')
}

plot_data = new_df 

ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  theme_bw()

#ggsave('./Write-up/Figures/speaker_noise_01_listener_1.png')
```

### p_noise = 0.05, prior_noise = 0.01

```{r warning = F, message = F}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 1/20 #listener noise
prior_prob_noise = 1/100 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.05.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.05.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.05.csv')
}

plot_data = new_df 

ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram() +
  geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  theme_bw()

```

### p_noise = 0.01, prior_noise = 0.05

```{r warning = F, message = F}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 1/100 #listener noise
prior_prob_noise = 1/20 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.05_listener_.01.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.05_listener_.01.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.05_listener_.01.csv')
}


plot_data = new_df 

ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram() +
  geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  theme_bw()

```

So clearly we get regularization when the listener assumes greater noise than the speaker produces. Let's try a few simulations where the difference is small:

### p_noise = 0.01, prior_noise = 0.005

```{r warning = F, message = F}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.01 #listener noise
prior_prob_noise = 0.005 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.005.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.005.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.005.csv')
}


plot_data = new_df 

p_lower_reg = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  theme_bw()

```

### p_noise = 0.01, prior_noise = 0.009

```{r warning = F, message = F}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.01 #listener noise
prior_prob_noise = 0.009 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.009.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.009.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.009.csv')
}


plot_data = new_df 

ngen_500_2 = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  theme_bw()
```

What if we have the same noise level but 1000 generations? (Maybe it takes longer to get to the stable generation with such a small difference between speaker and listener noise levels.

```{r warning = F, message = F}
n_gen = 1000 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.01 #listener noise
prior_prob_noise = 0.009 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.009_ngen1000.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.009_ngen1000.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.009_ngen1000.csv')
}


plot_data = new_df 

ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  theme_bw()

```

### p_noise = 0.01, prior_noise = 0.0075

```{r}
n_gen = 500 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.01 #listener noise
prior_prob_noise = 0.0075 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.0075.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.0075.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.0075.csv')
}


plot_data = new_df 

ngen_500 = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0.5, 1)) +
  theme_bw()
```

### 2000 generations

Since it seems like these eventually regularize, let's confirm this by trying a simulation with 2000 generations.

```{r warning = F, message = F}
n_gen = 2000 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.01 #listener noise
prior_prob_noise = 0.009 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.009_ngen1000.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.009_ngen2000.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.009_ngen2000.csv')
}


plot_data = new_df 

ngen_2000_2 = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0.5, 1)) +
  theme_bw()
```

```{r warning = F, message = F}
n_gen = 2000 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.01 #listener noise
prior_prob_noise = 0.0075 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.0075_ngen2000.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.0075_ngen2000.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.0075_ngen2000.csv')
}


plot_data = new_df 

ngen_2000 = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0.5, 1)) +
  theme_bw()

ggarrange(ngen_500_2, ngen_500, ngen_2000_2, ngen_2000, nrow = 2, ncol = 2)

#ggsave('./Write-up/Figures/fasterSlowerReg.png')
```

```{r}
n_gen = 2000 #number of generations
n_sims = 1000 #number of chains
p_theta = 0.5 #initial p_theta value

nu = 10
p_noise = 0.01 #listener noise
prior_prob_noise = 0.0075 #speaker noise

prior_mu = 0.6

df = as.data.frame(c(50, 100, 500, 1000, 10000))

no_cores = availableCores() - 1
plan(multisession, workers = no_cores)

file_path = '../Sims Data/speaker_.01_listener_.0075_ngen2000.csv'

if (!file.exists(file_path)) {


new_df = pmap_dfr(df, ~data.frame(iterated_learning(n_gen = n_gen, n_sims = n_sims, p_theta = 0.5, N = ..1, prior_mu = prior_mu, nu = nu, p_noise = p_noise, prior_prob_noise = prior_prob_noise, last_gen_only = T), 'Overall Frequency' = ..1))

write_csv(new_df, '../Sims Data/speaker_.01_listener_.0075_ngen2000.csv')

} else {
new_df = read_csv('../Sims Data/speaker_.01_listener_.0075_ngen2000.csv')
}


plot_data = new_df 

ngen_2000 = ggplot(data = plot_data, aes(x = posterior_mu, y = ..density..)) +
  geom_histogram(color = 'black', fill = 'white') +
  #geom_density() +
  facet_grid(generation ~ Overall.Frequency) +
  xlab(expression(mu[posterior]))+
  ylab('Density') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  theme_bw()
```
